Dispersion Distance and the Matter Distribution of the Universe in Dispersion Space

### 2015-07-29 writeup/referee_response.txt

This was chosen to be roughly the value for star forming galaxies at these
redshifts as measured by the WiggleZ survey. The assumption here is that FRBs
might trace star formation, although this is entirely speculative. It also
provides some contrast with the electrons, since some terms are proportional to
b_e - b_f. This is now stated in the text.

> Also, it would be good to discuss briefly (perhaps
> with a new plot) the dependence of the paper’s conclusions on the
> value of b_f. In principle, variations of b_f could cause interesting
> effects, since b_f affects some of the terms but not others in Eq.
> (22). My guess is that the effects won’t be too big, since Fig. 2
> shows that the integral term is the dominant contribution, but it’d be
> nice to numerically confirm this. If anything, it would be helpful to
> know that the predictions are robust to uncertainties in b_f.

This was a very helpful comment and led us to include a b_f=0.7 line in
Figure 3. We conclude that the sensitivity is indeed dominated by the
integral term and that a survey with 10000 events is unlikely to distinguish
any signature of the intrinsic clustering.

> 
> ii) To plot Fig. 3, the authors summed over pairs of radial bins.
> However, this hides some of the interesting statistical properties of
> the signal. For example, I would expect the cross power spectrum of
> two widely separated radial bins to be much lower than the cross power
> of two nearby bins. I would like to see some quantification of these
> radial correlations, as it would give readers a better sense for how
> the signal-to-noise scales with bin size, number of bins, etc.

Figure 2 is intended to give a hint of this structure. The integral term and
the cross term are fairly insensitive to radial separation since any structure
in the radial integral will correlate. The radial dependence is actually
dominated by A(\chi).  As shown in the figure, the local term drops rapidly at
higher separations. We would very much like to show more curves in the style of
Figure 2 but space constrains preclude this.

We've added a brief statement about this in the discussion.
We hope that the reviewer finds this satisfactory.

> 
> Finally, just an extremely minor pet peeve of mine: can we replace the
> error bars in Fig. 3 with a band around the curve? Real data would
> have scatter, and the error bars wouldn’t be centered on the “truth”.

Our error bars depend on \ell binning and thus a band is inappropriate.
We have switched to filled boxes.  This is consistent with the style in 
Scott Dodelson's text book.

> 
> Thank you to the authors for an interesting read!

### 2015-07-29 writeup/dm_space_distortions.tex
referee response.

For plotting the dispersion-space power spectra we use a toy model for the electron and
FRB clustering that assumes both are biased tracers of the dark matter:
\emph{i.e.}~$P_{xy}(k)=b_x b_y P(k)$.  We
calculate the time dependant dark-matter power spectrum using {\tt CAMB}, using
the integrated {\tt HALOFIT} to approximate the non-linear evolution. We assume
that the electron bias is $b_e=1$, which should be true on large scales due to
the approximate conservation of free electrons. The bias of FRB sources is
unconstrained by data and we choose it to be $b_f=1.3$. The coefficient
$A(\chi)$ depends on the density of detected FRB sources, $\bar n_f(\chi)$.
We use a simple model where all events above a fixed flux are detected,
the intrinsic luminosity function has the form from
\citet{1976ApJ...203..297S} with index $-1$, and events at the cut off of the
luminosity function and at
a radial distance of $\chi=2350\,{\rm Mpc}/h$ are at the flux detection threshold.
This source density is shown in Figure~\ref{f:n_f} with the resulting dispersion-space
power spectrum in Figure~\ref{f:terms}.

@@

@@

Perhaps, more concerning is the dispersion from the source's immediate environment
as well as the halo in which it resides. These electrons will presumably be
clustered around the source on scales far smaller than the survey
resolution and are thus not accounted for in the
$\delta_e$--$\delta_f$
correlations included in the above formalism.  This will have a mean
contribution, common to all sources and which may be epoch dependant, that will
modify the mean dispersion--distance relation. This will need to be calibrated in some
way or else fit with nuisance parameters during parameter estimation.
There will also be a stochastic piece which varies from source to source.  This will
limit the precision at which $\chi_s$ can be measured, thus limiting the
resolution of dispersion space density maps in the radial direction.
We note that one-point statistics from the same dataset 
should provide empirical information about these properties independent
of their physical origin.

@@

We consider the sensitivity of a survey with a total of $10000$ dispersion
measures from FRB events distributed uniformly over half the sky.
To give an idea of the overall sensitivity of the survey, we collapse
$C^{ss}_{\ell,ij}$ to a single function of $\ell$ by taking a weighted average
of all pairs of radial bins over the
range $500\,{\rm Mpc}/h$ to $3500\,{\rm Mpc}/h$.
For weights we use the signal over noise-squared
at $\ell=100$, which maximizes the signal to noise at that $\ell$ and
is near optimal at other multipoles.
The result is shown
in Figure~\ref{f:sensitivity}.

\begin{figure}
    \includegraphics[scale=0.42]{figures/sensitivity}
    \caption{
        \label{f:sensitivity}
        Sensitivity of a survey with $10000$ dispersion measures distributed
        uniformly over half the sky.  Plotted is the cross-correlation
        power spectrum weighted averaged over all pairs of radial bins. Weights
        are chosen to maximize signal to noise at $\ell=100$.
    }
\end{figure}

@@

We have proposed using cosmological FRB sources as standard pings to trace the 3D
large-scale structure of the Universe using dispersion distance. It can be seen in
Figure~\ref{f:terms} that the signal is dominated by inhomogeneities in the
electron density along the different lines-of-sight,
inducing mis-estimates of radial
distance and thus apparent clustering. These dispersion-space distortions could
be used
to study the distribution of free electrons in the Universe on large-scales.
There have been other proposals to use FRBs in cosmology 
\citep{2014ApJ...780L..33M,2014MNRAS.443L..11D,
2014ApJ...783L..35D,2014PhRvD..89j7303Z,2014ApJ...788..189G}, however these
schemes have either not been three dimensional
or required externally measured redshifts
for the sources.

We have
shown in Figure~\ref{f:sensitivity}
that a survey detecting $10000$ fast radio bursts of cosmological
origin could detect the
clustering signal. This is in rough
approximation to what might be achieved in a modest amount of time by the Canadian Hydrogen Intensity
Mapping Experiment (CHIME)\footnote{\url{http://chime.phas.ubc.ca/}}, the
Square Kilometer Array\footnote{\url{http://www.skatelescope.org/}},
or other upcoming wide-field telescopes.
Looking forward, dispersion-space
clustering could become a precision probe of large-scale structure should
observational factors turn out to be favourable compared to spectroscopic
surveys. The baryon acoustic oscillation feature can be clearly seen in
Figure~\ref{f:terms} in the cross term and both
local-term curves. It may also be possible to extract the BAO feature
from the integral
term using tomography.
This however requires that the statistics of the contribution to dispersion from
the source's environment and host halo to be well understood or small compared to the
contribution from the line-of-sight.

>>>>

For plotting the dispersion-space power spectra we use a toy model for the electron and
FRB clustering that assumes both are biased tracers of the dark matter:
\emph{i.e.}~$P_{xy}(k)=b_x b_y P(k)$.  We
calculate the time dependant dark-matter power spectrum using {\tt CAMB}, using
the integrated {\tt HALOFIT} to approximate the non-linear evolution. We assume
that the electron bias is $b_e=1$, which should be true on large scales due to
the approximate conservation of free electrons. The bias of FRB sources is
unconstrained by data and we choose it to be $b_f=1.3$\blue{, roughly
the value for star forming galaxies at the relevant
redshifts \citep{2011MNRAS.415.2876B} and providing reasonable contrast to the
electrons}.
The coefficient
$A(\chi)$ depends on the density of detected FRB sources, $\bar n_f(\chi)$.
We use a simple model where all events above a fixed flux are detected,
the intrinsic luminosity function has the form from
\citet{1976ApJ...203..297S} with index $-1$, and events at the cut off of the
luminosity function and at
a radial distance of $\chi=2350\,{\rm Mpc}/h$ are at the flux detection threshold.
This source density is shown in Figure~\ref{f:n_f} with the resulting dispersion-space
power spectrum in Figure~\ref{f:terms}.

@@

\blue{
The amplitude of the local and cross terms depend on the difference
between the electron bias and highly uncertain FRB bias
and the sign of the cross term depends on
which of these is larger. However, since the
integral term dominates for most radial distance pairs, our final sensitivity
estimates are largely unaffected by this uncertainty (as shown below). In a
high-precision survey the sensitivity of the cross and local terms to the
difference in the biases may help break measurement degeneracies between terms.
}

@@

Perhaps more concerning is the dispersion from the source's immediate environment
as well as the halo in which it resides. These electrons will presumably be
clustered around the source on scales far smaller than the survey
resolution and are thus not accounted for in the
$\delta_e$--$\delta_f$
correlations included in the above formalism.  This will have a mean
contribution, common to all sources and which may be epoch dependant, that will
modify the mean dispersion--distance relation. This will need to be calibrated in some
way or else fit with nuisance parameters during parameter estimation.
There will also be a stochastic piece which varies from source to source.  This will
limit the precision at which $\chi_s$ can be measured, thus limiting the
resolution of dispersion space density maps in the radial direction.
We note that one-point statistics from the same dataset 
should provide empirical information about these properties independent
of their physical origin.

@@

We consider the sensitivity of a survey with a total of $10000$ dispersion
measures from FRB events distributed uniformly over half the sky.
To give an idea of the overall sensitivity of the survey, we collapse
$C^{ss}_{\ell,ij}$ to a single function of $\ell$ by taking a weighted average
of all pairs of radial bins over the
range $500\,{\rm Mpc}/h$ to $3500\,{\rm Mpc}/h$.
For weights we use the signal over noise-squared
at $\ell=100$, which maximizes the signal to noise at that $\ell$ and
is near optimal at other multipoles.
\blue{
To determine if such a survey would be
able to distinguish between models of FRB clustering we plot the same quantity
for $b_f=0.7$.
}
The result is shown
in Figure~\ref{f:sensitivity}.

\begin{figure}
    \includegraphics[scale=0.42]{figures/sensitivity}
    \caption{
        \label{f:sensitivity}
        Sensitivity of a survey with $10000$ dispersion measures distributed
        uniformly over half the sky.  Plotted is the cross-correlation
        power spectrum weighted averaged over all pairs of radial bins. Weights
        are chosen to maximize signal to noise at $\ell=100$ for the $b_f=1.3$
        case.
    }
\end{figure}

@@

We have proposed using cosmological FRB sources as standard pings to trace the 3D
large-scale structure of the Universe using dispersion distance. It can be seen in
Figure~\ref{f:terms} that the signal is dominated by inhomogeneities in the
electron density along the different lines-of-sight,
inducing mis-estimates of radial
distance and thus apparent clustering.
\blue{
Because it is the line-of-sight integrated structure that correlates, the
signal is fairly insensitive to radial distance and radial separation of
slices, with these dependencies being dominated by the coefficient $A(\chi)$.
}
These dispersion-space distortions could
be used
to study the distribution of free electrons in the Universe on large-scales.
There have been other proposals to use FRBs for cosmology 
\citep{2014ApJ...780L..33M,2014MNRAS.443L..11D,
2014ApJ...783L..35D,2014PhRvD..89j7303Z,2014ApJ...788..189G}, however these
schemes have either not been three dimensional
or required externally measured redshifts
for the sources.

We have
shown in Figure~\ref{f:sensitivity}
that a survey detecting $10000$ fast radio bursts of cosmological
origin could detect the
clustering signal\blue{, however such a survey would be mostly sensitive to the
integral term and thus could not distinguish between FRB clustering models}.
This is in rough
approximation to what might be achieved in a reasonable amount of time by the
Canadian Hydrogen Intensity
Mapping Experiment (CHIME)\footnote{\url{http://chime.phas.ubc.ca/}}, the
Square Kilometer Array\footnote{\url{http://www.skatelescope.org/}},
or other upcoming wide-field telescopes.
Looking forward, dispersion-space
clustering could become a precision probe of large-scale structure should
observational factors turn out to be favourable compared to spectroscopic
surveys. The baryon acoustic oscillation feature can be clearly seen in
Figure~\ref{f:terms} in the cross term and both
local-term curves. It may also be possible to extract the BAO feature
from the integral
term using tomography.
This however requires that the statistics of the contribution to dispersion from
the source's environment and host halo to be well understood or small compared to the
contribution from the line-of-sight.
